---
title: "Connection Wizard"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Connection Wizard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r config, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

The QBMS connection wizard provides an interactive way to configure a QBMS
connection before retrieving data from a breeding-management system. It can be
used in three common contexts:

1. Directly from the R console or an R script.
2. Embedded inside a Shiny application.
3. Embedded inside an interactive R Markdown document that uses `runtime: shiny`.

The wizard has two interfaces:

- `qbms_wizard()` launches the wizard as a Shiny gadget. Use this from the R
  console or from a regular R script.
- `qbms_wizard_ui()` and `qbms_wizard_server()` expose the wizard as a Shiny
  module. Use this inside Shiny applications and interactive R Markdown
  documents.

## Option 1: Using the Wizard from the R Console

Use `qbms_wizard()` when working interactively from the R console or from a
regular R script. The wizard opens in the RStudio Viewer pane when available, or
in the default web browser. The R session waits until the user completes or
cancels the wizard.

```{r console-example}
library(QBMS)

result <- qbms_wizard()

if (is.null(result)) {
  message("Wizard was cancelled. No connection was configured.")
} else {
  message("Connection configured successfully.")
  message("Engine: ", result$engine)
  message("URL: ", result$url)

  print(result$selections)

  # The QBMS session is now configured, so QBMS data-retrieval
  # functions can be called directly. The exact functions available
  # depend on the selected engine and server.
  study_data <- get_study_data()
}
```

This approach is suitable for one-off interactive work, exploratory scripts, and
manual testing of a QBMS connection. It is not suitable for Shiny applications
or R Markdown documents running with `runtime: shiny`.

## Option 2: Embedding the Wizard in a Shiny Application

Inside a Shiny application, use the module interface. Add `qbms_wizard_ui()` to
the UI and call `qbms_wizard_server()` in the server function using the same
module ID.

The server function returns a reactive expression. Before the wizard is
completed, this reactive value is `NULL`. After completion, it contains the
connection result.

```{r shiny-app-example}
library(shiny)
library(QBMS)

ui <- fluidPage(
  titlePanel("QBMS Data Explorer"),

  sidebarLayout(
    sidebarPanel(
      qbms_wizard_ui("connection_wizard")
    ),

    mainPanel(
      h4("Connection result"),
      verbatimTextOutput("connection_result"),
      hr(),
      h4("Data preview"),
      tableOutput("data_preview")
    )
  )
)

server <- function(input, output, session) {
  wizard_result <- qbms_wizard_server("connection_wizard")

  output$connection_result <- renderPrint({
    result <- wizard_result()

    if (is.null(result)) {
      cat("Complete the wizard to configure the connection.")
    } else {
      cat("Connection configured.\n\n")
      cat("Engine:", result$engine, "\n")
      cat("URL:", result$url, "\n\n")
      cat("Selections:\n")
      print(result$selections)
    }
  })

  output$data_preview <- renderTable({
    result <- wizard_result()
    req(result)

    tryCatch(
      {
        # Replace this placeholder with the required QBMS call, for example:
        # head(get_study_data())
        data.frame(
          message = "Connection configured. Replace this with a QBMS data call.",
          engine = result$engine,
          url = result$url
        )
      },
      error = function(e) {
        data.frame(error = conditionMessage(e))
      }
    )
  })
}

shinyApp(ui = ui, server = server)
```

This pattern is the recommended approach when the connection wizard is part of a
larger Shiny application. Use `shiny::req()` or an equivalent guard before
calling QBMS data-retrieval functions that depend on the configured connection.

## Option 3: Embedding the Wizard in an Interactive R Markdown Document

Interactive R Markdown documents run inside a Shiny session when they use
`runtime: shiny`. For this reason, they should also use the module interface,
not `qbms_wizard()`.

The following is a minimal R Markdown document that embeds the connection
wizard and displays the result after completion.

````{verbatim}
---
title: "QBMS Connection Wizard"
output: html_document
runtime: shiny
---

```{r setup, include = FALSE}
library(QBMS)
```

## Configure the QBMS Connection

Complete the wizard below to configure the QBMS connection.

```{r wizard-ui, echo = FALSE}
qbms_wizard_ui("connection_wizard")
```

```{r wizard-server, context = "server"}
wizard_result <- qbms_wizard_server("connection_wizard")

output$wizard_result <- renderPrint({
  result <- wizard_result()

  if (is.null(result)) {
    cat("Complete the wizard above to configure the connection.")
  } else {
    cat("Connection configured.\n\n")
    cat("Engine:", result$engine, "\n")
    cat("URL:", result$url, "\n\n")
    cat("Selections:\n")
    print(result$selections)
  }
})
```

## Connection Result

```{r wizard-output, echo = FALSE}
verbatimTextOutput("wizard_result")
```
````

This approach is useful when preparing interactive reports or teaching material
where users need to configure a connection and then continue with live QBMS data
retrieval in the same document.

## Choosing the Right Interface

| Context | Recommended interface | Notes |
|---|---|---|
| R console or regular R script | `qbms_wizard()` | Opens an interactive gadget and returns the result directly. |
| Shiny application | `qbms_wizard_ui()` + `qbms_wizard_server()` | Use the module result as a reactive value. |
| R Markdown with `runtime: shiny` | `qbms_wizard_ui()` + `qbms_wizard_server()` | Use a UI chunk and a server-context chunk. |

> _**Notification:** Do not call `qbms_wizard()` from inside an already running Shiny session,
because it uses `shiny::runGadget()`. Use the module interface instead._

## Result Object

When the wizard is completed successfully, the returned object can be used to
inspect the selected connection settings and continue with QBMS data retrieval.
Typical fields include:

- `engine`: the selected QBMS engine;
- `url`: the server URL; and
- `selections`: the user selections made during the wizard.

Always check for `NULL` before using the result, because `NULL` indicates that
the wizard was cancelled or has not yet been completed.

> _**Development Note:** 
When testing the package from source, developers may load the local development
version with `devtools::load_all(".")`. Package vignettes, however, should use
`library(QBMS)` so that the vignette reflects how end users load the package._
