---
title: "QBMS Connection Wizard - R Markdown Example"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# remotes::install_github("icarda/QBMS")
# library(QBMS)
devtools::load_all(".")
```

## Using the QBMS Connection Wizard in R Markdown

This document demonstrates embedding the QBMS Connection Wizard in an
interactive R Markdown document. The document uses `runtime: shiny` to enable
Shiny interactivity.

**Important:** Since R Markdown with `runtime: shiny` runs inside a Shiny
session, we use the **module interface** (`qbms_wizard_ui` / `qbms_wizard_server`)
rather than `qbms_wizard()` (which would fail because `runGadget` cannot nest
inside a running Shiny app).

### Connection Wizard

Complete the steps below to configure your QBMS connection:

```{r wizard, echo=FALSE}
# Embed the wizard module UI
qbms_wizard_ui("my_wizard")
```

```{r wizard_server, context="server"}
# Activate the wizard module server
wizard_result <- qbms_wizard_server("my_wizard")

# Render the result below once completed
output$wizard_output <- renderPrint({
  result <- wizard_result()
  if (is.null(result)) {
    cat("Complete the wizard above to configure your connection.")
  } else {
    cat("Connection configured!\n\n")
    cat("Engine:", result$engine, "\n")
    cat("URL:", result$url, "\n")
    cat("\nSelections:\n")
    print(result$selections)
  }
})
```

### Result

```{r result_display, echo=FALSE}
verbatimTextOutput("wizard_output")
```
